---
title: "Tutorial 4 - Basic Data Manipulation"
tutorial:
  id: "com.example.tutorials.datamanipulation"
  version: 0.5.0
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to manipulate and wrangle your data.
---

```{r setup, include=FALSE, cache=FALSE}
library(learnr)
library(tidyverse) 
library(lubridate)
tutorial_options(exercise.timelimit = 60) # prevent overly complex code from running
knitr::opts_chunk$set(echo = FALSE)
```

```{r data-import, echo=FALSE, message=FALSE, warning=FALSE}
pinksalmonratio <- read_csv("../../data/OceanAK_PinkSalmonRatio_2000-2020.csv") %>%
  rename("catchdate" = "Catch Date",
         "statweek" = "Stat Week",
         "percentmale" = "Percent Males",	
         "samplesize" = "Sample Size",
         "weight_lbs" = "Fish Weight Pounds",	
         "num_fish_weighed" = "Number of Fish Weighed") %>%
  mutate(catchdate = ymd(as.POSIXct(catchdate, format = "%m/%d/%Y", tz = "US/Alaska")),
         Gear = as.factor(Gear),
         District = as.factor(District)) %>%
  dplyr::select(-Harvest)

```

## The PIPE

### Pipe - Exercise 1

*One shortcut that comes with dplyr is the pipe operator. This is written as *`%>%`*.  The pipe passes the results of the object/function on the left to the function on the right.*
*In practice, this is often employed by starting with an existing dataframe, then passing this to a function*
*E.g., dataframename %>% function()*

Using pipe syntax, re-write the code below to use a pipe:

```{r say-hello-to-your-little-friend, exercise=TRUE, exercise.eval=TRUE}
head(pinksalmonratio)
```

```{r say-hello-to-your-little-friend-solution}
pinksalmonratio %>% head()
```

### Pipe - Exercise 2

*Any function that requires a* `data=` *argument can just add the pipe in front to pass an item to it.* 

Similar to before, re-write the code below using `%>%` to pass `pinksalmonratio` to `str()`:

```{r pipe-str, exercise=TRUE, exercise.lines = 5}
str(pinksalmonratio)

```

```{r pipe-str-solution}
pinksalmonratio %>% str()

```


## Filtering

### Filter - Exercise 1

*To subset or remove rows of data that we do / don't want, we use the dplyr function* `filter()` *. We can use operators to keep (==), or remove (!=) rows that meet the conditions we set.*

Filter the data to only keep data from 2019:

```{r filter-2019, exercise=TRUE, exercise.eval=TRUE, exercise.lines = 5}
pinksalmonratio
```

```{r filter-2019-solution}
pinksalmonratio %>% filter(Year == 2019)
```


### Filter - Exercise 2

*We can also use multiple criteria*

Filter the data to only keep data from 2019 and where `statweek` is greater than 30:

```{r filter-2019mult, exercise=TRUE, exercise.eval=TRUE, exercise.lines = 5}
pinksalmonratio %>% filter(Year == 2019)
```

```{r filter-2019mult-solution}
pinksalmonratio %>% filter(Year == 2019, statweek > 30)
```


### Filter - Exercise 3

*Often values of NAs are of interest, either to highlight missing data or to remove NAs from further analysis*

Filter the data to only show our `NA` values from `weight_lbs`:
```{r filter-weightnas, exercise=TRUE, exercise.lines = 5}
pinksalmonratio %>% filter(is.na())
```

```{r filter-weightnas-solution}
pinksalmonratio %>% filter(is.na(weight_lbs))
```


Filter the data to the subset where District is 101 and that excludes `weight_lbs` values that are `NA`:
```{r filter-weight_notnas, exercise=TRUE, exercise.lines = 5}
pinksalmonratio %>% filter(District == "", !is.na())
```

```{r filter-weight_notnas-solution}
pinksalmonratio %>% filter(District == "101", !is.na(weight_lbs))
```


### Filter - Exercise 3 - Review
Did you 154 rows? Great job!  
Notice how you needed to put 101 in quotes?? This is because it isn't a number anymore, it's a factor! 


### Filter - Exercise 4
*We can also use operators such as * `|` *(OR),* `&` *(AND)*

Filter the data to the subset where District is 101 OR 102:
```{r filter-operator, exercise=TRUE, exercise.lines = 5}
pinksalmonratio %>% filter(District == "" | District == "")
```

```{r filter-operator-solution}
pinksalmonratio %>% filter(District == "101" | District == "102")
```


### Filter - Exercise 5

*Finally, we can also use the operator* `%in%` *as a shortcut to filter in a list*

Filter the data to only show Districts from the list of `c("101", "102", "103")` using `%in%`:
```{r filter-in, exercise=TRUE, exercise.lines = 5}
pinksalmonratio %>% filter(District)
```

```{r filter-in-solution}
pinksalmonratio %>% filter(District %in% c("101", "102", "103"))
```



## Selecting and renaming columns

### Rename - Exercise 1
*Often column names will need to be renamed using * `rename()`  
*For `rename()`, put the column names in quotations.  

Rename column `District` to `District_fact`. If more help is needed, learn more using `?rename`
```{r rename, exercise=TRUE, exercise.lines = 5}
pinksalmonratio %>% rename()
```

```{r rename-solution}
pinksalmonratio %>% rename("District_fact" = "District")
```


### Select - Exercise 1
*Another tool that we use is to keep or drop columns using * `select()`.
*Note that you do* __not__ *need quotes around columns for *`select()`. 

Only keep columns `Year` and `percentmale`. 
```{r selectkeep, exercise=TRUE, exercise.lines = 5}
pinksalmonratio %>% select()
```

```{r selectkeep-solution}
pinksalmonratio %>% select(Year, percentmale)
```


### Select - Exercise 2
`select()` *can also be used to drop a column by adding a negative sign in front of a column.*

Drop columns `samplesize`, `weight_lbs`, `num_fish_weighed`. 
```{r selectdrop, exercise=TRUE, exercise.lines = 5, exercise.eval=FALSE}
pinksalmonratio %>% select(-)
```

```{r selectdrop-solution}
pinksalmonratio %>% select(-samplesize, -weight_lbs, -num_fish_weighed)
```



## Mutate

### Mutate - Exercise 1

*Adding a new column is also quite easy in dplyr using * `mutate()`  
*You do not need column names in quotes.*  

Add a new column called `Comment` where every value says, "Great sampling!"
```{r mutate-comment, exercise=TRUE, exercise.lines = 5, exercise.eval=FALSE}
pinksalmonratio %>% mutate()
```

```{r mutate-comment-solution}
pinksalmonratio %>% mutate(Comment = "Great sampling!")
```


### Mutate - Exercise 2

Add a new column called `Gear_District` of both the `Gear` and `District` columns.
For this exercise, use `paste0()` to paste values together. Play around until it looks correct. 
```{r mutatenew, exercise=TRUE, exercise.lines = 5, exercise.eval=FALSE}
pinksalmonratio %>% mutate(Gear_District = )
```

```{r mutatenew-solution}
pinksalmonratio %>% mutate(Gear_District = paste0(Gear, District))
```


### Mutate - Exercise 3

For year 2018, find the average weight per fish by dividing column `weight_lbs` by `num_fish_weighed`
```{r mutate-weight, exercise=TRUE, exercise.lines = 5, exercise.eval=FALSE}
pinksalmonratio %>% filter(Year == 2018) %>% mutate(meanweight = )
```

```{r mutate-weight-solution}
pinksalmonratio %>% filter(Year == 2018) %>% mutate(meanweight = weight_lbs / num_fish_weighed)
```


### Mutate - Exercise 4
*An if else statement can also be used to conditionally set a value.* 
*For this exercise, `ifelse()` has 3 arguments. The first argument is the criteria, the second is what to put if the criteria is true, and the third is what to put if it is false.*

Add a new column where if the statweek is 27 or under it is labeled "early", otherwise label it "late"
```{r mutate-if, exercise=TRUE, exercise.lines = 5, exercise.eval=FALSE}
pinksalmonratio %>% mutate(earlylateseason = )
```

```{r mutate-if-solution}
pinksalmonratio %>% mutate(earlylateseason = ifelse(statweek <= 27, "early", "late"))
```


### Mutate - Exercise 5
*Mutate is often used to turn a column into a factor.* 
*The function * `as.factor()` *can be combined with mutate to change a factor into a column. You should be aware of the implications of this!*

Change column 'Year' to be a factor
```{r mutate-factor, exercise=TRUE, exercise.lines = 5, exercise.eval=FALSE}
pinksalmonratio %>% mutate(  = as.factor())


```

```{r mutate-factor-solution}
pinksalmonratio %>% mutate(Year = as.factor(Year))
```


### Mutate - Exercise 6
*Now let's turn a factor back into a number. This can be trickier than it first appears*  
*Review the results of the first line. Because District was a factor, each factor was simply one of 16 "levels", one for each district. To return it to a number, first convert it to a character,* __then__ *to a number.* 

Turn column District into a numeric value based on its District value. 
```{r mutate-factorreturn, exercise=TRUE, exercise.lines = 5, exercise.eval=FALSE}
pinksalmonratio %>% mutate(District = as.numeric(District))

#pinksalmonratio %>% mutate( = as.numeric(as.character()))


```

```{r mutate-factorreturn-solution}
pinksalmonratio %>% mutate(District = as.numeric(as.character(District)))
```









## Quiz Part 1

*Here are a few questions to review the code concepts we've covered. These scores are not submitted anywhere so they're only for your self-guidance. If you have a question about any of these answers, please ask.*  

Answer the following:  

```{r data-manip-quiz1}
quiz(
  question("Why should you do data manipulation in R rather than Excel?",
    answer("You have a record of what changes were implemented", correct = TRUE),
    answer("Can be easily repeated every time you output data from OceanAK / data is updated", correct = TRUE),
    answer("Easier & faster for many tasks", correct = TRUE),
    answer("Because Justin said to")
  ),
  question("How would you filter dataframe 'pinksalmonratio' to only include values from 2018 in column 'Year'?",
    answer("pinksalmonratio %>% filter(Year == 2018)", correct = TRUE),
    answer("pinksalmonratio %>% filter(Year = 2018)"),
    answer("pinksalmonratio$Year %>% filter(2018)"),
    answer("filter(pinksalmonratio, Year == 2018)", correct = TRUE)
  ),
    question("Multiple filtering criteria: How would you filter dataframe 'pinksalmonratio' to only include values from 2018 in column 'Year' as well as 'Pink salmon' for column 'Species'?",
    answer("pinksalmonratio %>% filter(Year == 2018, 'Pink salmon')"),
    answer("pinksalmonratio %>% filter(Year == 2018, Species == 'Pink salmon')", correct = TRUE),
    answer("pinksalmonratio %>% filter(Year == 2018 & Species == 'Pink salmon')", correct = TRUE),
    answer("pinksalmonratio %>% filter(year == 2018, species == 'Pink salmon')"),
    answer("pinksalmonratio %>% filter(Year == 2018 | Species == 'Pink salmon')")
  ),
    question("Select the correct code to rename column 'Year' to 'harvestyear'.",
    answer("pinksalmonratio %>% rename('Year' = 'harvestyear')"),
    answer("pinksalmonratio %>% rename('year' == 'harvestyear')"),
    answer("pinksalmonratio %>% rename('harvestyear' = 'Year')", correct = TRUE)
  )
)
```



### Quiz 1 Review
That was tricky! Here are a few reminders:  
 - When checking equality in a `filter()` statement you'll need a double equals (==)  
 - The ampersand (&) means AND which adds more and more specific criteria  
 - The vertical bar (|) mean OR which allows alternate criteria  
 - Using a comma in `filter()` is a shortcut to allow alternate criteria (similar to |)  
 - When renaming, the new column name is specified first, then the old column name. See `?rename` for more info  
 

## Quiz Part 2

*Here are a few more quiz questions. *  

Answer the following:  

```{r data-manip-quiz2}
quiz(
  question("Keep only the columns named 'Year' and 'percentmale' from 'pinksalmonratio' ",
    answer("pinksalmonratio %>% select(Year, percentmale)", correct = TRUE),
    answer("pinksalmonratio %>% keepcolumns('Year', 'percentmale')"),
    answer("pinksalmonratio %>% dplyr::select(Year, percentmale)", correct = TRUE),
    answer("pinksalmonratio %>% select(-Year, -percentmale)")
  ),
  question("Drop the column 'Year' from 'pinksalmonratio'. ",
    answer("pinksalmonratio %% select(Year)"),
    answer("pinksalmonratio %>% donotselect(Year)"),
    answer("pinksalmonratio %>% select(-Year)", correct = TRUE)
  ),
  question("How would you add a new column (called 'avweight') that divides column 'weight_lbs' by 'num_fish_weighed'? ",
    answer("pinksalmonratio %>% Mutate(avweight == weight_lbs / num_fish_weighed)"),
    answer("pinksalmonratio %>% mutate(avweight = weight_lbs / num_fish_weighed)", correct = TRUE),
    answer("mutate(pinksalmonratio, avweight = weight_lbs / num_fish_weighed)", correct = TRUE)
  ),
  question("Using ifelse statements are helpful for conditional criteria. How would you add a new column called 'highlowpercent' based on criteria from column 'percentmale'? ",

    answer("ifelse(pinksalmonratio$percentmale >80, 'High', 'low')"),
        answer("pinksalmonratio %>% mutate(highlowpercent = ifelse(percentmale >80, 'High Percent Males', 'Low'))", correct = TRUE),
    answer("pinksalmonratio %>% mutate(ifelse(percentmale >80, 'High Percent Males', 'Low'))"),
    answer("pinksalmonratio %>% mutate(highlowpercent = ifelse(percentmale, 80, 'High', 'Low'))")
  )
)
```


### Quiz 2 Review
Great job! Let's review a few points quickly:    
 - In `mutate` and `select` we aren't required to quotes around the column names (but you can if wanted)  
 - For `ifelse` statements, the syntax is `ifelse(criteria, valueiftrue, valueiffalse)` 


### COMPLETE
That's all, congratulations on getting through this lesson!
