---
title: "Tutorial 2 - Basics"
tutorial:
  id: "com.example.tutorials.basics"
  version: 0.0.2
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn the basics of how R works.
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(lubridate)


tutorial_options(exercise.timelimit = 60) # prevent overly complex code from running
knitr::opts_chunk$set(echo = FALSE)
```

```{r loadpkg, echo=FALSE, message=FALSE, warning=FALSE}

# Now load in some data that we'll use later
tannercrab <- read_csv("../../data/OceanAK_CrabSurvey_2018-2020_Specimens.csv") %>%
  rename("Count" = "Number Of Specimens",
         "Width" = "Width Millimeters",
         "Chela" = "Chela Height Millimeters",
         "SurveyStartDate" = "Trip Start Date") %>%
  mutate(SurveyStartDate = ymd(as.POSIXct(SurveyStartDate, format = "%m/%d/%Y", tz = "US/Alaska"))) %>%
  dplyr::select(Year, SurveyStartDate, Location, Species, Sex, Width, Chela, Count) %>%
  filter(Species == "Bairdi tanner crab", Count >= 1, !is.na(Width)) # Keep only tanners & remove 0s
```

# ADF&G learnr Tutorial
## Basic R Programming 

### Exercise 1 {data-allow-skip=TRUE}

*Use this space to work through code to understand the concepts we've gone over. Go slow and try to get the correct answers. This is not graded and is not a race, use this space to familiarize yourself with how R works. Once you are finished, just close out of this before opening a new tutorial.*
*This mini-console is running R, however clicking "Run Code" evaluates everything written in there; in RStudio, we run through it line-by-line.*

Write the R code required to add two plus two:

```{r twoplus, exercise=TRUE, exercise.lines = 5}

```

```{r twoplus-solution}
2+2
```

### Exercise 2

*Here's an exercise with some prepopulated code.*

Complete the code to concatenate numbers 1 through 5 together to make a vector called "myvector" then print "myvector" to the screen. 

```{r make-vector, exercise=TRUE, exercise.lines = 5}
myvector <- (, , , , )
```


```{r make-vector-solution}
myvector <- c(1, 2, 3, 4, 5)
myvector
```

<div id="make-vector-hint">
**Hint:** c(1, 2, 3, 4 ,5)
</div>

### Exercise 2 - Review
Woohoo! You can see that the code `myvector <- c(1, 2, 3, 4, 5)` set myvector to be 1 through 5, then when you type `myvector` it prints it to the screen.



## Dataframes

### Dataframes - Exercise 1

*In this exercise we will use a dataframe that I've already loaded for you. This dataframe is called "tannercrab".*

Modify the following code to show only the top 10 rows:

```{r print-limit, exercise = TRUE, exercise.eval=TRUE, exercise.lines = 5}
tannercrab
```

```{r print-limit-solution}
head(tannercrab, n=10)
```

<div id="print-limit-hint">
**Hint:** Modify head() 
</div>


### Dataframes - Exercise 2

*Let's work with just the width of the crab*
  
Using the `$` operator, select from the `tannercrab` dataframe just the column named "Width"  

```{r columnWidth, exercise = TRUE, exercise.lines = 5}

```

```{r columnWidth-solution}
tannercrab$Width
```

<div id="columnWidth-hint">
  **Hint:** tannercrab$
</div>
  
  
  
### Dataframes - Exercise 3
  
  *Now let's do some very basic analysis.*

What is the mean width of all tanner crab?

```{r meanwidth, exercise = TRUE, exercise.lines = 5}
mean()
```

```{r meanwidth-solution}
mean(tannercrab$Width)
```
<div id="meanwidth-hint">
**Hint:** mean(tannercrab$Width)
</div>


### Dataframes - Exercise 3 - Review
Did you get 126.9? GREAT JOB!


### Dataframes - Exercise 4

*Let's now look at the mean Chela width.*
  
Take the mean of column "Chela". (NOTE! There are many NAs in here!)

```{r chelamean, exercise = TRUE, exercise.lines = 5}
mean(tannercrab$Chela)
```

```{r chelamean-solution}
mean(tannercrab$Chela, na.rm = TRUE)
```

<div id="chelamean-hint">
  **Hint:** Do you need an argument in mean() of na.rm = TRUE?
</div>

### Dataframes - Exercise 4 Solution

*Solution: * `mean(tannercrab$Chela, na.rm = TRUE)`  
*As you saw, the * `NA` *caused issues when we took the mean. The solution was to add the argument * `na.rm = TRUE` *to the * `mean()` *function.*
 *This gives us a mean chela width of about 25.3 mm*




### Dataframes - Exercise 5

*BUG HUNTER*
  
Whoops! Looks like someone made an error in their code. Fix this to run.

```{r tanner-bug, exercise = TRUE, exercise.lines = 5}
Mean(Tannercrab$chela, Na.rm = TRUE)
```

```{r tanner-bug-solution}
mean(tannercrab$Chela, na.rm = TRUE)
```

<div id="tanner-bug-hint">
  **Hint:** There are four capitalization issues
</div>



## Quiz

*Let's review some of the concepts we've gone over.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz-basics}
quiz(
  question("Which function would you use for taking the mean?",
    answer("average()"),
    answer("tools"),
    answer("mean()", correct = TRUE),
    answer("codetools")
  ),
  question("Choose how to select the 7th column named 'Chela' of a dataframe named 'tannercrab':",
    answer("tannercrab[,7]", correct = TRUE),
    answer("tannercrab[,column7]"),
    answer("tannercrab[7,]"),
    answer("tannercrab$Chela", correct = TRUE),
    answer("tannercrab[,'Chela']", correct = TRUE)
  )
)
```

