---
title: "Tutorial 2 - Basics"
tutorial:
  id: "com.example.tutorials.basics"
  version: 0.1.0
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn the basics of how R works.
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(lubridate)


tutorial_options(exercise.timelimit = 60) # prevent overly complex code from running
knitr::opts_chunk$set(echo = FALSE)
```

```{r loadpkg, echo=FALSE, message=FALSE, warning=FALSE}

# Now load in some data that we'll use later
tannercrab <- read_csv("../../data/OceanAK_CrabSurvey_2018-2020_Specimens.csv") %>%
  rename("Count" = "Number Of Specimens",
         "Width" = "Width Millimeters",
         "Chela" = "Chela Height Millimeters",
         "SurveyStartDate" = "Trip Start Date") %>%
  mutate(SurveyStartDate = ymd(as.POSIXct(SurveyStartDate, format = "%m/%d/%Y", tz = "US/Alaska"))) %>%
  dplyr::select(Year, SurveyStartDate, Location, Species, Sex, Width, Chela, Count) %>%
  filter(Species == "Bairdi tanner crab", Count >= 1, !is.na(Width)) # Keep only tanners & remove 0s
```

# ADF&G learnr Tutorial
## Basic R Programming 

### Exercise 1 {data-allow-skip=TRUE}

*Use this space to work through code to understand the concepts we've gone over. Go slow and try to get the correct answers. This is not graded and is not a race, use this space to familiarize yourself with how R works. Once you are finished, just close out of this before opening a new tutorial.*
*This mini-console is running R, however clicking "Run Code" evaluates everything written in there; in RStudio, we run through it line-by-line.*

Write the R code required to add two plus two:

```{r twoplus, exercise=TRUE, exercise.lines = 5}

```

```{r twoplus-solution}
2+2
```

### Exercise 2

*Here's an exercise with some prepopulated code.*

Complete the code to concatenate numbers 1 through 5 together to make a vector called "myvector" then print "myvector" to the screen. 

```{r make-vector, exercise=TRUE, exercise.lines = 5}
myvector <- (, , , , )
```


```{r make-vector-solution}
myvector <- c(1, 2, 3, 4, 5)
myvector
```

<div id="make-vector-hint">
**Hint:** c(1, 2, 3, 4 ,5)
</div>

### Exercise 2 - Review
Woohoo! You can see that the code `myvector <- c(1, 2, 3, 4, 5)` set myvector to be 1 through 5, then when you type `myvector` it prints it to the screen.



## Dataframes

### Dataframes - Exercise 1

*In this exercise we will use a dataframe that I've already loaded for you. This dataframe is called "tannercrab".*

Modify the following code to show only the top 10 rows:

```{r print-limit, exercise = TRUE, exercise.eval=TRUE, exercise.lines = 5}
tannercrab
```

```{r print-limit-solution}
head(tannercrab, n=10)
```

<div id="print-limit-hint">
**Hint:** Modify head() 
</div>


### Dataframes - Exercise 2

*Let's work with just the width of the crab*
  
Using the `$` operator, select from the `tannercrab` dataframe just the column named "Width"  

```{r columnWidth, exercise = TRUE, exercise.lines = 5}

```

```{r columnWidth-solution}
tannercrab$Width
```

<div id="columnWidth-hint">
  **Hint:** tannercrab$
</div>
  
  
  
### Dataframes - Exercise 3
  
*Now let's do some very basic analysis.*

What is the mean width of all tanner crab?

```{r meanwidth, exercise = TRUE, exercise.lines = 5}
mean()
```

```{r meanwidth-solution}
mean(tannercrab$Width)
```
<div id="meanwidth-hint">
**Hint:** mean(tannercrab$Width)
</div>


### Dataframes - Exercise 3 - Review
Did you get 126.9? GREAT JOB!


### Dataframes - Exercise 4

*Let's now look at the mean Chela width.*
  
Take the mean of column "Chela". (NOTE! There are many NAs in here!)

```{r chelamean, exercise = TRUE, exercise.lines = 5}
mean(tannercrab$Chela)
```

```{r chelamean-solution}
mean(tannercrab$Chela, na.rm = TRUE)
```

<div id="chelamean-hint">
  **Hint:** Do you need an argument in mean() of na.rm = TRUE?
</div>

### Dataframes - Exercise 4 Solution

*Solution: * `mean(tannercrab$Chela, na.rm = TRUE)`  
*As you saw, the * `NA` *caused issues when we took the mean. The solution was to add the argument * `na.rm = TRUE` *to the * `mean()` *function.*
 *This gives us a mean chela width of about 25.3 mm*



### Dataframes - Exercise 5

*BUG HUNTER*
  
Whoops! Looks like someone made an error in their code. Fix this to run.

```{r bug-tanner, exercise = TRUE, exercise.lines = 5}
Mean(Tannercrab$chela, Na.rm = TRUE)
```

```{r bug-tanner-solution}
mean(tannercrab$Chela, na.rm = TRUE)
```

<div id="bug-tanner-hint">
  **Hint:** There are four capitalization issues
</div>



### Dataframes - Exercise 6

**BUG HUNTER Part 2**
  
What the $%@#! There are STILL more errors to fix. Fix this to run.

```{r bug-meanvec, exercise = TRUE, exercise.lines = 5}
mean(as.vector(c(1 2, 3, 4, 5, NA) na.rm = TRUE)
```

```{r bug-meanvec-solution}
mean(as.vector(c(1, 2, 3, 4, 5, NA)), na.rm = TRUE)
```

<div id="bug-meanvec-hint">
  **Hint:** There are three comma / parenthesis issues
</div>



### Dataframes - Exercise 7

**BUG HUNTER Part 3**
  
Oh No! There's another line with some errors in the code. Fix this to run.

```{r bug-vector, exercise = TRUE, exercise.eval=FALSE, exercise.lines = 5}
c(1 2, 3, 4, 5, NA))
```

```{r bug-vector-solution}
c(1, 2, 3, 4, 5, NA)
```

<div id="bug-vector-hint">
  **Hint:** There are two comma / parenthesis issues
</div>



## Quiz Part 1 

*Let's review some of the concepts we've gone over.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz-basics1}
quiz(
  question("Which function would you use for taking the mean?",
    answer("average()"),
    answer("tools"),
    answer("mean()", correct = TRUE),
    answer("codetools")
  ),
  question("Choose how to select the 7th column named 'Chela' of a dataframe named 'tannercrab':",
    answer("tannercrab[,7]", correct = TRUE),
    answer("tannercrab[,column7]"),
    answer("tannercrab[7,]"),
    answer("tannercrab$Chela", correct = TRUE),
    answer("tannercrab[,'Chela']", correct = TRUE)
  ),
  question("How would you set variable 'fishcolor' to be equal to red, blue, and green?",
    answer("fishcolor <- red, blue, green "),
    answer("fishcolor = c('red', 'blue', 'green') ", correct = TRUE),
    answer("fishcolor <- c('red', 'blue', 'green')", correct = TRUE),
    answer("fishcolor <- c(red, blue, green)"),
    answer("'fishcolor' <- c('red, blue, green')")
  ),
  question("How would you get help and learn more about a function named cbind()?",
    answer("help(cbind)", correct = TRUE),
    answer("Search cbind in help panel on bottom-right of RStudio", correct = TRUE),
    answer("Google search 'r cbind()'", correct = TRUE),
    answer("Call 867-5309 and ask for cbind"),
    answer("?cbind", correct = TRUE)
  ),
  question("Select the only line that does NOT have an issue",
    answer("mean(c(3, 4, 5, 6)"),
    answer("mean(2, 3, 4, 5)"),
    answer("mean(c(3, 4, 5, 6))", correct = TRUE),
    answer("mean(x"),
    answer("mean((x)")
  ),
  question("What is the expected outcome of this code: mean(c(2, 2, 4, 6, 6, NA, 2, 6, 5, 3, 4))",
    answer("NA", correct = TRUE),
    answer("4")
  )

)
```


## Quiz Part 2

*Great job! Now let's try some similar problems.*

Here are few more questions that build on what we just learned:

```{r quiz-basics2}
quiz(
  question("Which of the following is a concatenated group of numbers?",
    answer("makelist(1, 2, 3)"),
    answer("str(1, 2, 3)"),
    answer("listed(1, 2, 3)"),
    answer("c(1, 2, 3)", correct = TRUE)
  ),
  question("Select column 'Length' that is the 2nd column in dataframe 'groundfish':",
    answer("groundfish$Length", correct = TRUE),
    answer("groundfish[,'Length']", correct = TRUE),
    answer("groundfish[,column2]"),
    answer("groundfish[2,]"),
    answer("groundfish[,2]", correct = TRUE),
    answer("groundfish(col=2)")
  ),
  question("While taking the mean, your answer is returned as NA. What should you add to the code?",
    answer("completecase = FALSE"),
    answer("na.rm = TRUE", correct = TRUE),
    answer("pleaseremoveNAs")
  ),
  question("Oh no! Your console shows a '+' and appears to be waiting for you. What caused this and how to you get it back to normal?",
    answer("Cause: Code is running. Solution: Wait. Wait some more. Wait again?"),
    answer("Cause: Incorrect voodoo chant. Solution: Burn computer"),
    answer("Cause: Wrong function name. Solution: Never make a mistake ever again"),
    answer("Cause: Incomplete code ending. Solution: Press Escape", correct = TRUE)
  ),
  question("Which of these is a good variable name",
    answer("x"),
    answer("sablefishcatch2019", correct = TRUE),
    answer("mean"),
    answer("vector"),
    answer("dat"),
    answer("1sablefish")
  )
)

```


